### 2025-04-22 scripts/kb-data-transformations.rmd

---
title: "R Notebook"
output: html_notebook
---

##  eva_pubyear_comparison.csv

```{r}
library(readr)
library(dplyr)
library(tidyr)

# Rename column to target values
data <- read_csv("data/eva_pubyear_comparison.csv") |>
  rename(
    publisher = num_articles_publisher,
    openalex = num_articles_openalex
  )

# Pivot data into wider format
wider_data <- data |>
  select(-diff) |>
  pivot_longer(
    cols = c(publisher, openalex),
    names_to = "source",
    values_to = "value"
  )

write_csv(wider_data, "data/eva_pubyear_comparison_pivoted.csv")
```

##  jura_groels_alle_zeitschriften.csv
```{r}
library(readr)
library(dplyr)
library(tidyr)

data <- read_csv("data/jura_groels_alle_zeitschriften.csv") |>
  select(-zeitschrift, -in_wos_emerging_sources_index)

# Pivot data into wider format
wider_data <- data |>
  mutate(across(.cols = c("issn_list", "wos_source_title", "wos_issn", "scopus_source_id", "scp_issn", "openalex_source_id", "openalex_issn", "crossref_issn"),
                .fns = function(x) ifelse(is.na(x), 0, 1))) |>
  pivot_longer(
    cols = c("issn_list", "wos_source_title", "wos_issn", "scopus_source_id", "scp_issn", "openalex_source_id", "openalex_issn", "crossref_issn"),
    names_to = "source",
    values_to = "value"
  )

write_csv(wider_data, "data/jura_groels_alle_zeitschriften_pivoted.csv")
```

##  juristenzeitung_authors_affiliations_openalex.csv

```{r}
library(readr)
library(dplyr)
library(tidyr)

# "institution_id","raw_author_name","author_id","cleaned_author_name","cleaned_institution_name","min_pubyear","max_pubyear","publ_cnt","example_id"

data <- read_csv("data/juristenzeitung_authors_affiliations_openalex.csv") |>
  select(author_id, cleaned_author_name, institution_id, cleaned_institution_name) |>
  group_by(author_id)  |>
  filter(!is.na(institution_id) | row_number() == min(row_number()))  |>
  slice(1)

```

```{r}
library(dplyr)
library(readr)

impact_factor_df <- read_csv("data/eva_journal_impact_factors.csv") |>
  group_by(abk) |>
  summarize(journal_impact_factor_avg = mean(journal_impact_factor, na.rm = TRUE))

general_journals_df <- read_csv("data/jura_groels_tabelle1_allg_zeitschriften.csv") |> select(abk, qualitaet_gewichtet)
specialized_journals_df <- bind_rows(read_csv("data/jura_groels_tabelle2_fachspez_zeitschriften.csv") |> select(abk, qualitaet_gewichtet))

```

```{r}
library(dplyr)
library(readr)

df <- read_csv("data/juristenzeitung_authors_affiliations_openalex.csv", 
  locale = locale(encoding = "UTF-8")) |>
  mutate(author_id = as.numeric(factor(author_id)),
         institution_id = as.numeric(factor(institution_id)))

vertices_authors <- df |>
  select(id = author_id, label = cleaned_author_name) |>
  filter(!is.na(id)) |>
  mutate(type = "author") |>
  distinct()

vertices_institutions <- df |>
  select(id = institution_id, label = cleaned_institution_name) |>
  filter(!is.na(id)) |>
  mutate(type = "institution") |>
  distinct()

vertices <- bind_rows(vertices_authors, vertices_institutions)

# Assemble the nodes information in the target format
nodes_info <- paste("*nodes\nId;Label;Type", 
                    paste(vertices$id, vertices$label, vertices$type, sep = ";", collapse = "\n"), sep = "\n")

edges <- df |> 
  filter(!is.na(author_id) & !is.na(institution_id)) |>
  select(from = author_id, to = institution_id)

# Assemble the edges information in the target format
edges_info <- paste("*edges\nId;Id", 
                    paste(edges$from, edges$to, sep = ";", collapse = "\n"), sep = "\n")

# Concatenate nodes and edges information
graph_info <- paste(nodes_info, edges_info, sep = "\n\n")

# save data to file
writeLines(graph_info, "data/juristenzeitung_authors_affiliations_openalex_network.csv", 
  useBytes = TRUE)

```